---
title: "Cards against humanity paper notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stargazer)
library(mice)
library(brms)
library(gridExtra)
```

### Read the data

```{r}
df_raw = read_csv('/Users/georgeberry/Desktop/CAH/cah_data_files/201710-CAH_PulseOfTheNation_Raw.csv')
```

### Clean the data

```{r}
dep_var = 'civil_war'
indep_var = 'lost_friends'
control_vars = c(
  'income',
  'gender',
  'age',
  'age_range',
  'pol_affil',
  'race',
  'educ',
  'kale',
  'hunting',
  'nationalist',
  'love_amer'
)
all_vars = c(dep_var, indep_var, control_vars)
```

```{r}
colnames(df_raw) = c(
  'income',
  'gender',
  'age',
  'age_range',
  'pol_affil',
  'approve',
  'educ',
  'blank1',
  'race',
  'blank2',
  'nationalist',
  'rep_nationalist',
  'blank3',
  'love_amer',
  'dem_love_amer',
  'blank4',
  'help_poor',
  'rep_help_poor',
  'white_racist',
  'dem_white_racist',
  'lost_friends',
  'civil_war',
  'hunting',
  'kale',
  'the_rock',
  'vader_or_trump'
)

df = df_raw %>%
  select(all_of(all_vars)) %>%
  mutate(across(everything(), ~ifelse(. == 'DK/REF', NA, .))) %>%
  mutate(
    civil_war = ifelse(civil_war == 'Likely', 1, 0),
    log_income = log1p(income),
    #income=income/1000,
  ) %>%
  mutate_if(is.character, as.factor) %>%
  mutate(
    gender=relevel(gender, 'Female'),
    age_range=relevel(age_range, '35-44'),
    pol_affil=relevel(pol_affil, 'Independent'),
    race=relevel(race, 'White'),
    educ=relevel(educ, 'High school'),
    kale=relevel(kale, 'No'),
    hunting=relevel(hunting, 'No'),
    nationalist=relevel(nationalist, 'Disagree'),
    love_amer=relevel(love_amer, 'No'),
    lost_friends=relevel(lost_friends, 'No')
  )
```


### Missingness structure

```{r}
colSums(is.na(df))
```

```{r}
colSums(
  is.na(
    df %>%
      select(-log_income)
  )
) -> missing_table
missing_table = as_tibble(as.data.frame(missing_table), rownames = 'Demographic') %>%
  rename(`Missing rows`=missing_table)
stargazer(missing_table, summary=F, rownames=F)
```

### Examining lost friends

Overall

```{r}
mean(df$lost_friends == 'Yes', na.rm=T)
```

Break down by demo

```{r}
df %>%
  select(-income, -log_income, -age) %>%
  filter(!is.na(lost_friends)) %>%
  mutate(
    civil_war = ifelse(civil_war == 1, 'Yes', 'No')
  ) %>%
  pivot_longer(c(civil_war, gender:love_amer)) %>%
  mutate(
    lost_friends = ifelse(lost_friends == 'Yes', 1, 0),
  ) %>%
  filter(!is.na(value), value != 'Other') %>%
  group_by(name, value) %>%
  summarize(
    se = sd(lost_friends, na.rm=T) / sqrt(n()),
    lost_friends = mean(lost_friends),
    N=n()
  ) %>%
  mutate(
    demo = paste0(name, '—', value)
  ) %>%
  arrange(demo) -> lost_friends_demos
lost_friends_demos
```

```{r}
plot_lost_friends = ggplot(lost_friends_demos) +
  geom_point() +
  aes(x=reorder(demo, lost_friends), y=lost_friends) +
  geom_hline(yintercept=0, linetype='dotdash') +
  geom_hline(yintercept=1, linetype='dotdash') +
  ylim(c(0, 1)) +
  labs(
    title='Fraction who lost a friend in\nresponse to 2016 election',
    y='Lost friends',
    x='Demographic'
  ) +
  coord_flip() +
  theme_minimal()
plot_lost_friends
```


Friendship loss is high for the 25-34 age group and for non-Republicans

### Examining perceptions of civil war

Overall

```{r}
mean(df$civil_war, na.rm=T)
```

Wow! 1/3 of Americans think a civil war is "Likely". That's a pretty high level of fear.

Break down by demo

```{r}
df %>%
  select(-income, -log_income, -age) %>%
  filter(!is.na(civil_war)) %>%
  pivot_longer(lost_friends:love_amer) %>%
  filter(!is.na(value), value != 'Other') %>%
  group_by(name, value) %>%
  summarize(
    se = sd(civil_war, na.rm=T) / sqrt(n()),
    civil_war = mean(civil_war),
    N=n()
  ) %>%
  mutate(
    demo = paste0(name, '—', value)
  ) %>%
  arrange(demo) -> civil_war_demos
civil_war_demos
```

Super interesting: the people who lost friends are different than the people who think a civil war is likely in the oneway! For instance the education pattern is reversed.


```{r}
plot_civil_war = ggplot(civil_war_demos) +
  geom_point() +
  aes(x=reorder(demo, civil_war), y=civil_war) +
  geom_hline(yintercept=0, linetype='dotdash') +
  geom_hline(yintercept=1, linetype='dotdash') +
  ylim(c(0, 1)) +
  labs(
    title='Fraction who said a civil war\nwas "Likely" in the next decade',
    y='Said civil war "Likely"',
    x='Demographic'
  ) +
  coord_flip() +
  theme_minimal()
plot_civil_war
```


### Impute and regress

Note the following

- We do not use the DV `civil_war` as part of imputation
- We drop rows in the regression where the main independent variable `lost_friends` is missing

drop cases where DV is null, impute, and regress

```{r}
df_no_dv = df %>% select(-civil_war)

imp_no_dv = mice(df_no_dv, m=5, maxit=50, meth='pmm', seed=500, printFlag=FALSE)
df_imp_no_dv = complete(imp_no_dv, 1)

df2 = df_imp_no_dv
df2$civil_war = df$civil_war
df2 = df2[!is.na(df$lost_friends),]
```

After accounting for both demos and some lifestyle questions, people who lost friends as a result of the 2016 election are 12 percentage points more likely to view a civil war as "Likely". That's large!

```{r}
mod2 = lm(
  civil_war ~ lost_friends + log1p(income) + gender + age_range + pol_affil + race + educ + kale + hunting + nationalist + love_amer,
  data=df2
)
summary(mod2)
```

We are imputing 70% of income so let's check a model without income to make sure we get similar results

```{r}
df_no_inc = df %>% select(-civil_war, -income, -log_income)

imp_no_inc = mice(df_no_inc, m=5, maxit=50, meth='pmm', seed=500, printFlag=FALSE)
df_imp_no_inc = complete(imp_no_inc, 1)

df3 = df_imp_no_inc
df3$civil_war = df$civil_war
df3 = df3[!is.na(df$lost_friends),]
```

Interesting, the lifestyle variables matter here!

```{r}
mod3 = lm(
  civil_war ~ lost_friends + gender + age_range + pol_affil + race + educ + kale + hunting + nationalist + love_amer,
  data=df3
)
summary(mod3)
```

And just out of curiosity, what's the raw regression say?

```{r}
mod2_raw = lm(
  civil_war ~ lost_friends,
  data=df2
)
summary(mod2_raw)
```

And finally, let's do imputation excluding the indep var (`lost_frends`) from imputation. This is a check that matters if we interpret losing friends causally.

```{r}
df_no_dviv = df %>% select(-civil_war, -lost_friends)

imp_no_dviv = mice(df_no_dviv, m=5, maxit=50, meth='pmm', seed=500, printFlag=FALSE)
df_imp_no_dviv = complete(imp_no_dviv, 1)

df4 = df_imp_no_dviv
df4$civil_war = df$civil_war
df4$lost_friends = df$lost_friends
df4 = df4[!is.na(df$lost_friends),]
```

```{r}
mod4 = lm(
  civil_war ~ lost_friends + log1p(income) + gender + age_range + pol_affil + race + educ + kale + hunting + nationalist + love_amer,
  data=df4
)
summary(mod4)
```

Just curious: what if we drop the lifestyle questions or do only lifestyle

```{r}
summary(
  lm(
    civil_war ~ lost_friends + kale + hunting + nationalist + love_amer,
    data=df2
  )
)
```

```{r}
summary(
  lm(
    civil_war ~ lost_friends + log1p(income) + gender + age_range + pol_affil + race + educ,
    data=df2
  )
)
```

### Now do logits or else people will (wrongly) complain

```{r}
mod2_logit = glm(
  civil_war ~ lost_friends + log1p(income) + gender + age_range + pol_affil + race + educ + kale + hunting + nationalist + love_amer,
  data=df2,
  family=binomial
)
summary(mod2_logit)
```
```{r}
mod3_logit = glm(
  civil_war ~ lost_friends + gender + age_range + pol_affil + race + educ + kale + hunting + nationalist + love_amer,
  data=df3, 
  family=binomial
)
summary(mod3_logit)
```

```{r}
mod4_logit = glm(
  civil_war ~ lost_friends + log1p(income) + gender + age_range + pol_affil + race + educ + kale + hunting + nationalist + love_amer,
  data=df4,
  family=binomial
)
summary(mod4_logit)
```
```{r}
mod2_raw_logit = glm(civil_war ~ lost_friends, family=binomial, data=df2)
summary(mod2_raw_logit)
```

Logistic regression gut check, we should get a similar probability to OLS

```{r}
df_yes = df2 %>%
  mutate(
    lost_friends='Yes'
  )
df_no = df2 %>%
  mutate(
    lost_friends='No'
  )

mean(predict(mod2_logit, newdata=df_yes, type='response') - predict(mod2_logit, newdata=df_no, type='response'))
```

Bam, it works!

### A simple Bayesian regression for the skeptical

```{r}
mod_brms = brm(
  civil_war ~ lost_friends + log1p(income) + gender + age_range + pol_affil + race + educ + kale + hunting + nationalist + love_amer,
  data=df2,
  family=bernoulli,
  refresh=0
)
summary(mod_brms)
```


### Make tables

Have to convert odds ratios se's like so: https://www.andrewheiss.com/blog/2016/04/25/convert-logistic-regression-standard-errors-to-odds-ratios-with-r/

```{r}
coef.mod2_raw_logit <- exp(mod2_raw_logit$coef)
ses.mod2_raw_logit <- sqrt(coef.mod2_raw_logit^2 * diag(vcov(mod2_raw_logit)))
pval.mod2_raw_logit <- summary(mod2_raw_logit)$coef[, 4]

coef.mod2_logit <- exp(mod2_logit$coef)
ses.mod2_logit <- sqrt(coef.mod2_logit^2 * diag(vcov(mod2_logit)))
pval.mod2_logit <- summary(mod2_logit)$coef[, 4]

coef.mod3_logit <- exp(mod3_logit$coef)
ses.mod3_logit <- sqrt(coef.mod3_logit^2 * diag(vcov(mod3_logit)))
pval.mod3_logit <- summary(mod3_logit)$coef[, 4]


library(stargazer)
#stargazer(model.count,coef=list(coef.vector))

stargazer(
  mod2_logit,
  mod3_logit, 
  mod2_raw_logit,
  coef=list(
    coef.mod2_logit,
    coef.mod3_logit,
    coef.mod2_raw_logit
  ),
  se=list(
    ses.mod2_logit,
    ses.mod3_logit,
    ses.mod2_raw_logit
  ),
  p=list(
    pval.mod2_logit,
    pval.mod3_logit,
    pval.mod2_raw_logit
  ),
  #type='html',
  column.labels=c(
    "Primary Model",
    "No Income Model",
    "Bivariate Model"
  ),
  no.space=TRUE,
  star.cutoffs = c(0.05, 0.01, 0.001)
  # report=('vc*p')
)
```

### Make plots

White, income

```{r}
# white, income

new_cah.df.a <- data.frame(
  income=seq(0, 250000, 10000),
  lost_friends='No',
  age_range = '35-44',
  gender = "Male",
  pol_affil = "Independent",
  educ = "College degree",
  race = "White",
  kale = 'Yes',
  hunting = 'No',
  nationalist = 'Disagree',
  love_amer = 'Yes'
)
new_cah.df.b <- data.frame(
  income=seq(0, 250000, 10000),
  lost_friends='Yes',
  age_range = '35-44',
  gender = "Male",
  pol_affil = "Independent",
  educ = "College degree",
  race = "White",
  kale = 'Yes',
  hunting = 'No',
  nationalist = 'Disagree',
  love_amer = 'Yes'
)
new_cah.df.c <- data.frame(
  income=seq(0, 250000, 10000),
  lost_friends='No',
  age_range = '35-44',
  gender = "Female",
  pol_affil = "Independent",
  educ = "College degree",
  race = "White",
  kale = 'Yes',
  hunting = 'No',
  nationalist = 'Disagree',
  love_amer = 'Yes'
)
new_cah.df.d <- data.frame(
  income=seq(0, 250000, 10000),
  lost_friends='Yes',
  age_range = '35-44',
  gender = "Female",
  pol_affil = "Independent",
  educ = "College degree",
  race = "White",
  kale = 'Yes',
  hunting = 'No',
  nationalist = 'Disagree',
  love_amer = 'Yes'
)
new_cah.df <- rbind(new_cah.df.a, new_cah.df.b, new_cah.df.c, new_cah.df.d)

new_cah.df$prediction <- predict(mod2_logit, new_cah.df, type='response')

plot_white_income <- ggplot(new_cah.df) +
  aes(
    x = income,
    y = prediction,
    linetype = as.factor(lost_friends),
    color = as.factor(gender)
  ) + 
  geom_smooth(se=FALSE) +
  ggtitle('Prediction for Likely/Unlikely Civil War \nGiven Lost Friend (White, College Degree)') + 
  xlab('Income') +
  ylab('Predicted Probability') +
  labs(linetype='Lost Friends?', color='Gender') +
  ylim(0,1) +
  theme_minimal() +  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))


plot_white_income
```

```{r}
# black, income

new_cah.df.a <- data.frame(
  income=seq(0, 250000, 10000),
  lost_friends='No',
  age_range = '35-44',
  gender = "Male",
  pol_affil = "Independent",
  educ = "High school",
  race = "Black",
  kale = 'Yes',
  hunting = 'No',
  nationalist = 'Disagree',
  love_amer = 'Yes'
)
new_cah.df.b <- data.frame(
  income=seq(0, 250000, 10000),
  lost_friends='Yes',
  age_range = '35-44',
  gender = "Male",
  pol_affil = "Independent",
  educ = "High school",
  race = "Black",
  kale = 'Yes',
  hunting = 'No',
  nationalist = 'Disagree',
  love_amer = 'Yes'
)
new_cah.df.c <- data.frame(
  income=seq(0, 250000, 10000),
  lost_friends='No',
  age_range = '35-44',
  gender = "Female",
  pol_affil = "Independent",
  educ = "High school",
  race = "Black",
  kale = 'Yes',
  hunting = 'No',
  nationalist = 'Disagree',
  love_amer = 'Yes'
)
new_cah.df.d <- data.frame(
  income=seq(0, 250000, 10000),
  lost_friends='Yes',
  age_range = '35-44',
  gender = "Female",
  pol_affil = "Independent",
  educ = "High school",
  race = "Black",
  kale = 'Yes',
  hunting = 'No',
  nationalist = 'Disagree',
  love_amer = 'Yes'
)
new_cah.df <- rbind(new_cah.df.a, new_cah.df.b, new_cah.df.c, new_cah.df.d)

new_cah.df$prediction <- predict(mod2_logit, new_cah.df, type='response')

plot_black_income <- ggplot(new_cah.df) +
  aes(
    x = income,
    y = prediction,
    linetype = as.factor(lost_friends),
    color = as.factor(gender)
  ) + 
  geom_smooth(se=FALSE) +
  ggtitle(
    'Prediction for Likely/Unlikely Civil War \nGiven Lost Friend (Black, High School)'
  ) + 
  xlab('Income') +
  ylab('Predicted Probability') +
  labs(linetype='Lost Friends?', color='Gender') +
  ylim(0,1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))

plot_black_income
```

### Combine plots

```{r}
grid.arrange(plot_white_income, plot_black_income, nrow=1)
```

```{r}
grid.arrange(plot_lost_friends, plot_civil_war, nrow=1)
```





### left to do

4. revise
